\singlespacing \vspace{-1.6cm} \LARGE Trade-offs among plant reproductive traits determine interactions with floral visitors

### 2022-01-31 Manuscript/Lanuza_et_al_2021_Ecol_Letters.tex
Edits to fit the max of 5000 words

Despite the astonishing diversity of floral structures among flowering plants (\protect\hyperlink{ref-barrett2002}{Barrett 2002}; \protect\hyperlink{ref-schiestl2013}{Schiestl \& Johnson 2013}) and their importance in shaping plant-pollinator interactions (\protect\hyperlink{ref-fenster2004}{Fenster \emph{et al.} 2004}; \protect\hyperlink{ref-dellinger2020}{Dellinger 2020}), a unified framework that \textbf{describes the major} plant reproductive trade-offs is currently lacking (\protect\hyperlink{ref-roddy2021}{Roddy \emph{et al.} 2021}). In addition, macroecological studies that investigate plant reproductive traits are scarce (\protect\hyperlink{ref-baude2016}{Baude \emph{et al.} 2016}; \protect\hyperlink{ref-munoz2016}{Munoz \emph{et al.} 2016}; \protect\hyperlink{ref-grossenbacher2017}{Grossenbacher \emph{et al.} 2017}; \protect\hyperlink{ref-moeller2017}{Moeller \emph{et al.} 2017}) and consequently, there is poor understanding of how reproductive traits drive interactions with floral visitors at large scales (\protect\hyperlink{ref-sargent2008}{Sargent \& Ackerly 2008}; \protect\hyperlink{ref-rech2016}{Rech \emph{et al.} 2016}; \protect\hyperlink{ref-salguero2016}{Salguero-Gómez \emph{et al.} 2016}; \protect\hyperlink{ref-ruger2018}{Rüger \emph{et al.} 2018}). Linking the plant's position in trait-space with the different pollinator groups could help to improve our understanding of plant-pollinator associations (\protect\hyperlink{ref-dehling2016}{Dehling \emph{et al.} 2016}). Further, there is increasing interest in understanding drivers of plant-pollinator interactions using trait-based approaches (\protect\hyperlink{ref-fenster2004}{Fenster \emph{et al.} 2004}; \protect\hyperlink{ref-rosas2014}{Rosas-Guerrero \emph{et al.} 2014}) and trait-matching analyses (\protect\hyperlink{ref-stang2009}{Stang \emph{et al.} 2009}; \protect\hyperlink{ref-bartomeus2016}{Bartomeus \emph{et al.} 2016}). \textbf{However, plant reproductive traits have been overlooked beyond highly specialised pollination systems (\protect\hyperlink{ref-dellinger2020}{Dellinger 2020}) and remains unclear how specific plant reproductive biology traits (e.g., mating or compatibility system) influence plant-pollinator associations (\protect\hyperlink{ref-tur2013}{Tur \emph{et al.} 2013}; \protect\hyperlink{ref-devaux2014}{Devaux \emph{et al.} 2014}).}

Species can optimise their fitness through various life-history traits, yet trade-offs among those traits constrain the range of potential strategies that a species can use. With the recent availability of large trait databases (e.g., TRY, \protect\hyperlink{ref-kattge2011}{Kattge \emph{et al.} 2011}; and COMPADRE, \protect\hyperlink{ref-salguero2015}{Salguero-Gómez \emph{et al.} 2015}), plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function (\protect\hyperlink{ref-diaz2016}{Díaz \emph{et al.} 2016}; \protect\hyperlink{ref-salguero2016}{Salguero-Gómez \emph{et al.} 2016}; \protect\hyperlink{ref-bruelheide2018}{Bruelheide \emph{et al.} 2018}; \protect\hyperlink{ref-carmona2021}{Carmona \emph{et al.} 2021}). However, most studies have focused on vegetative traits such as leaf (\protect\hyperlink{ref-wright2004}{Wright \emph{et al.} 2004}), wood (\protect\hyperlink{ref-chave2009}{Chave \emph{et al.} 2009}), or root (\protect\hyperlink{ref-laughlin2021}{Laughlin \emph{et al.} 2021}) trade-offs with little or no attention given to reproductive traits (\protect\hyperlink{ref-evojtko2020}{E-Vojtkó \emph{et al.} 2020}; \protect\hyperlink{ref-roddy2021}{Roddy \emph{et al.} 2021}) which are critical to plant life strategies that shape interactions with pollinators and ultimately determine plant reproductive success. For instance, short lived versus perennial species tend to have low versus high levels of outcrossing, respectively, (\protect\hyperlink{ref-barrett2003}{Barrett 2003}; \protect\hyperlink{ref-moeller2017}{Moeller \emph{et al.} 2017}) and outcrossing levels are positively correlated with flower size (\protect\hyperlink{ref-goodwillie2010}{Goodwillie \emph{et al.} 2010}). In addition, the presence of costly rewards (e.g., pollen or nectar) and showy flowers or floral displays can only be understood through consideration of plant species' reliance upon animal pollination (pollinator dependence) and its role in attracting pollinators (\protect\hyperlink{ref-ollerton2011}{Ollerton \emph{et al.} 2011}; \protect\hyperlink{ref-rodger2021}{Rodger \emph{et al.} 2021}). However, it is still unknown to what extent these different reproductive compromises determine plant-pollinator interactions.

Several studies have identified links between plant traits and plant-pollinator network properties (\protect\hyperlink{ref-bartomeus2013}{Bartomeus 2013}; \protect\hyperlink{ref-olito2015}{Olito \& Fox 2015}; \protect\hyperlink{ref-rowe2020}{Rowe \emph{et al.} 2020}). Moreover, plant traits can define species' network roles (e.g., specialists vs generalists; \protect\hyperlink{ref-lazaro2013}{Lázaro \emph{et al.} 2013}; \protect\hyperlink{ref-tur2013}{Tur \emph{et al.} 2013}). For example, plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators (\protect\hyperlink{ref-junker2013}{Junker \emph{et al.} 2013}; \protect\hyperlink{ref-coux2016}{Coux \emph{et al.} 2016}). Morphological matching between plant and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure (\protect\hyperlink{ref-stang2009}{Stang \emph{et al.} 2009}; \protect\hyperlink{ref-ibanez2012}{Ibanez 2012}). Remarkably, the combination of traits has shown to increase the predictive power of the network interactions (\protect\hyperlink{ref-eklof2013}{Eklöf \emph{et al.} 2013}). \textbf{Moreover, the position in the trait space can determine plant and pollinator species' functional roles (\protect\hyperlink{ref-dehling2016}{Dehling \emph{et al.} 2016}). Therefore, by considering the multidimensional reproductive trait space we could progress our understanding of plant-pollinator associations. Importantly, we know little if those patterns generally studied at the community level are representative of wider macroecological scales.}

@@

All species names, genera, families and orders were retrieved and standardized from the taxonomy data sources NCBI (\url{https://www.ncbi.nlm.nih.gov/taxonomy}) for plants and ITIS (\url{https://www.itis.gov/}) for pollinators, using the R package \emph{taxize} (\protect\hyperlink{ref-chamberlain2020}{Chamberlain \emph{et al.} 2020}) version \emph{0.9.99}. We filled the `not found' searches manually using \url{http://www.theplantlist.org/} and \url{http://www.mobot.org/} for plants and \url{http://www.catalogueoflife.org/} for floral visitors.

@@

We selected a total of 19 different functional traits that included both reproductive and vegetative traits (Table 1 and Supplementary Information). From these, 16 were reproductive traits (13 floral and 3 reproductive biology traits) and were selected based on their relevance to plant reproduction and data availability. Note that autonomous selfing was recorded quantitatively (fruit set) and qualitatively because this trait was rarely available in its numerical form. The 3 remaining traits, were vegetative traits that are commonly used to characterize the global spectrum of plant form and function and represent the fast-slow continuum of trait variation (e.g., short-lived versus perennial species). For each plant species, we undertook an extensive literature and online search across a wide range of resources (plant databases, online floras, books, journals and images). From a total of 30,120 cells (20 columns x 1506 species) we were able to fill 24,341 cells (80.8\% of the dataset, see Fig. S1 for missing values information for each trait).

@@

\caption{\label{tab:unnamed-chunk-1}\textbf{Table 1} Plant functional traits used in this study.}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{>{}cl>{}ccl}
\toprule
\multicolumn{2}{c}{\textbf{Quantitative traits}} & \multicolumn{3}{c}{\textbf{Categorical traits}} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-5}
\textbf{Type} & \textbf{Traits} & \textbf{Type} & \textbf{Traits} & \textbf{Categories}\\
\midrule
\textbf{Vegetative} & \makecell[c]{Plant height \\ (m)} & \textbf{Vegetative} & Lifepan & \makecell[l]{Short-lived \\ Perennial}\\
\addlinespace
\textbf{Floral} & \makecell[c]{Flower width \\ (mm)} & \textbf{Vegetative} & Life form & \makecell[l]{Herb \\ Shrub \\ Tree}\\
\addlinespace
\textbf{Floral} & \makecell[c]{Flower length \\ (mm)} & \textbf{Floral} & \makecell[c]{Flower \\ shape} & \makecell[l]{Brush \\ Campanulate \\ Capitulum \\ Open \\ Papilionaceous \\ Tube}\\
\addlinespace
\textbf{Floral} & \makecell[c]{Inflorescence width \\ (mm)} & \textbf{Floral} & \makecell[c]{Flower \\ symmetry} & \makecell[l]{Actinomorphic \\ Zygomorphic}\\
\addlinespace
\textbf{Floral} & \makecell[c]{Style length \\ (mm)} & \textbf{Floral} & Nectar & \makecell[l]{Presence \\ Absence}\\
\addlinespace
\textbf{Floral} & Ovules per flower & \textbf{\makecell[c]{Reproductive \\ biology}} & \makecell[c]{Compatibility \\ system} & \makecell[l]{Self-incompatible \\ Part. self-compatible \\ Self-compatible}\\
\addlinespace
\textbf{Floral} & Flowers per plant & \textbf{\makecell[c]{Reproductive \\ biology}} & \makecell[c]{Breeding \\ system} & \makecell[l]{Hermaphrodite \\ Monoecious \\ Dioecious}\\
\addlinespace
\textbf{Floral} & \makecell[c]{Nectar \\ ($\mu$l)} & \textbf{\makecell[c]{Reproductive \\ biology}} & \makecell[c]{Autonomous \\ selfing} & \makecell[l]{None \\ Low \\ Medium \\ High}\\
\addlinespace
\textbf{Floral} & \makecell[c]{Nectar \\ (mg)} & \textbf{} &  & \\
\addlinespace
\textbf{Floral} & \makecell[c]{Nectar concentration \\ ($\%$)} & \textbf{} &  & \\
\addlinespace
\textbf{Floral} & \makecell[c]{Pollen grains \\ per flower} & \textbf{} &  & \\
\addlinespace
\textbf{\makecell[c]{Reproductive \\ biology}} & \makecell[c]{Autonomous selfing \\ (fruit set)} & \textbf{} &  & \\
\bottomrule
\end{tabular}
\end{table}

@@

Trait missing values were imputed with the function \emph{missForest} (\protect\hyperlink{ref-stekhoven2012}{Stekhoven \& Bühlmann 2012}) which allows imputation of data sets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of \emph{missForest} (\protect\hyperlink{ref-penone2014}{Penone \emph{et al.} 2014}). To extract the eigenvectors, we used the function \emph{PVRdecomp} from the package \emph{PVR} (\protect\hyperlink{ref-santos2018}{Chamberlain \emph{et al.} 2018}) based on a previous conceptual framework that considers phylogenetic eigenvectors (\protect\hyperlink{ref-diniz-filho2012}{Diniz-Filho \emph{et al.} 2012}). However, we were unable to include nectar and pollen traits on the imputation process because of the high percentage of missing values (Fig. S1). Hence, the imputed dataset had 1,506 species, seven categorical and eight numerical variables and 5.79\% of missing values. Further, we conducted an additional imputation process on the subset of species with data for pollen per flower and microliters of nectar. This subset comprised 755 species, 8.01\% missing values and all traits but milligrams of nectar (\textasciitilde50\% of missing values) were included in the imputation process.

@@

We explored the trade-offs between different quantitative plant functional traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson's correlation = 0.72, \emph{P} \textless{} 0.01 and Pearson's correlation = 0.36, \emph{P} \textless{} 0.01), and thus we avoided overemphasizing flower size on the reproductive spectrum of trait variation. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the association of the different qualitative traits with the main axes of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers, we excluded values within the 2.5th--97.5th percentile range (\protect\hyperlink{ref-legendre2012}{Legendre \& Legendre 2012}), and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range. We performed the pPCA using the function \emph{phyl.pca} from the package \emph{phytools} version \emph{0.7-70} (\protect\hyperlink{ref-revell2012}{Revell 2012}) with the method lambda (\(\lambda\)) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation (\protect\hyperlink{ref-abdi2010}{Abdi \& Williams 2010}). Moreover, to corroborate that our imputation of missing values did not affect our results, we conducted a pPCA on the full dataset without missing values (Fig. S2). We found little difference between the explained variance with the imputed dataset (51.08\%) and the dataset without missing values (52.87\%). In addition, the loadings on each principal component had a similar contribution and correlation patterns, with the exception of plant height which showed slight variations between the imputed and non-imputed dataset. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

\textbf{Phylogenetic signal of traits}

We calculated the phylogenetic signal of the different quantitative traits on the imputed dataset with the full set of species (N = 1,506) with the package \emph{phytools} (\protect\hyperlink{ref-revell2012}{Revell 2012}) version \emph{0.7-70} and we used Pagel's \(\lambda\) as a measurement of the phylogenetic signal. However, for pollen and nectar traits, phylogenetic signal was calculated only on the subset of species that had quantitative information for these traits (N = 755).

@@

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. These networks were analysed in their qualitative (presence-absence) and quantitative (interaction frequency) form. First, we analysed the binary version of these weighted networks with presence-absence information that assumes equal weight across interactions. Second, we analysed the untransformed weighted networks with interaction frequency that accounts for the intensity of the interaction. Although floral visitors are not always pollinators and interaction frequency does not consider each pollinator species efficiency (\protect\hyperlink{ref-ballantyne2015}{Ballantyne \emph{et al.} 2015}), interaction frequency can provide valuable information of the contribution of floral visitors to pollination (\protect\hyperlink{ref-vazquez2005}{Vázquez \emph{et al.} 2005}, \protect\hyperlink{ref-vazquez2012}{2012}). In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across networks but this was accounted for in analyses by considering them in the random effects of the modelling process. All analyses were conducted in R version \emph{4.0.3}.

@@

We used Bayesian modelling (see below for details) to explore the effect of floral visitor groups and the main axes of trait variation (pPCA with imputed dataset) on both qualitative (presence/absence) and quantitative (visitation rate) floral interactions per plant species. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits of all groups, we also explored the presence-absence of interaction and visitation rate of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that \emph{Apis mellifera} was the floral visitor with the largest proportion of records counted (7.55\% of the total). This finding is consistent with previous research showing that \emph{A. mellifera} was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems (\protect\hyperlink{ref-hung2018}{Hung \emph{et al.} 2018}). Hence, to control for the effect of \emph{A. mellifera} on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of interaction and visitation rate excluding \emph{A. mellifera}. We found that \emph{A. mellifera}, was partly driving some of the observed trends on PC1 (Fig. S3). However, we did not detect major differences on PC2 and PC3.

>>>>

Despite the astonishing diversity of floral structures among flowering plants (\protect\hyperlink{ref-barrett2002}{Barrett 2002}; \protect\hyperlink{ref-schiestl2013}{Schiestl \& Johnson 2013}) and their importance in shaping plant-pollinator interactions (\protect\hyperlink{ref-fenster2004}{Fenster \emph{et al.} 2004}; \protect\hyperlink{ref-dellinger2020}{Dellinger 2020}), a unified framework that \textbf{describes the major} plant reproductive trade-offs is currently lacking (\protect\hyperlink{ref-roddy2021}{Roddy \emph{et al.} 2021}). In addition, macroecological studies that investigate plant reproductive traits are scarce (\protect\hyperlink{ref-baude2016}{Baude \emph{et al.} 2016}; \protect\hyperlink{ref-munoz2016}{Munoz \emph{et al.} 2016}; \protect\hyperlink{ref-grossenbacher2017}{Grossenbacher \emph{et al.} 2017}; \protect\hyperlink{ref-moeller2017}{Moeller \emph{et al.} 2017}) and consequently, there is poor understanding of how reproductive traits drive interactions with pollinators at large scales (\protect\hyperlink{ref-sargent2008}{Sargent \& Ackerly 2008}; \protect\hyperlink{ref-rech2016}{Rech \emph{et al.} 2016}; \protect\hyperlink{ref-salguero2016}{Salguero-Gómez \emph{et al.} 2016}; \protect\hyperlink{ref-ruger2018}{Rüger \emph{et al.} 2018}). Linking the plant's position in trait-space with the different pollinator groups could help to improve our understanding of plant-pollinator associations (\protect\hyperlink{ref-dehling2016}{Dehling \emph{et al.} 2016}). Further, there is increasing interest in understanding drivers of plant-pollinator interactions using trait-based approaches (\protect\hyperlink{ref-fenster2004}{Fenster \emph{et al.} 2004}; \protect\hyperlink{ref-rosas2014}{Rosas-Guerrero \emph{et al.} 2014}) and trait-matching analyses (\protect\hyperlink{ref-stang2009}{Stang \emph{et al.} 2009}; \protect\hyperlink{ref-bartomeus2016}{Bartomeus \emph{et al.} 2016}). \textbf{However, plant reproductive traits have been overlooked beyond highly specialised pollination systems (\protect\hyperlink{ref-dellinger2020}{Dellinger 2020}) and remains unclear how specific plant reproductive biology traits (e.g., mating or compatibility system) influence plant-pollinator associations (\protect\hyperlink{ref-tur2013}{Tur \emph{et al.} 2013}; \protect\hyperlink{ref-devaux2014}{Devaux \emph{et al.} 2014}).}

Species can optimise their fitness through various life-history traits, yet trade-offs among those traits constrain the range of potential strategies that a species can use. With the recent availability of large trait databases (e.g., TRY \protect\hyperlink{ref-kattge2011}{Kattge \emph{et al.} 2011}; and COMPADRE \protect\hyperlink{ref-salguero2015}{Salguero-Gómez \emph{et al.} 2015}), plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function (\protect\hyperlink{ref-diaz2016}{Díaz \emph{et al.} 2016}; \protect\hyperlink{ref-salguero2016}{Salguero-Gómez \emph{et al.} 2016}; \protect\hyperlink{ref-bruelheide2018}{Bruelheide \emph{et al.} 2018}; \protect\hyperlink{ref-carmona2021}{Carmona \emph{et al.} 2021}). However, most studies have focused on vegetative traits such as leaf (\protect\hyperlink{ref-wright2004}{Wright \emph{et al.} 2004}), wood (\protect\hyperlink{ref-chave2009}{Chave \emph{et al.} 2009}), or root (\protect\hyperlink{ref-laughlin2021}{Laughlin \emph{et al.} 2021}) trade-offs with little or no attention given to reproductive traits (\protect\hyperlink{ref-evojtko2020}{E-Vojtkó \emph{et al.} 2020}; \protect\hyperlink{ref-roddy2021}{Roddy \emph{et al.} 2021}) which are critical to plant life strategies that shape interactions with pollinators and ultimately determine plant reproductive success. For instance, short lived versus perennial species tend to have low versus high levels of outcrossing, respectively, (\protect\hyperlink{ref-barrett2003}{Barrett 2003}; \protect\hyperlink{ref-moeller2017}{Moeller \emph{et al.} 2017}) and outcrossing levels are positively correlated with flower size (\protect\hyperlink{ref-goodwillie2010}{Goodwillie \emph{et al.} 2010}). In addition, the presence of costly rewards (e.g., pollen or nectar) and showy flowers or floral displays can only be understood through consideration of plant species' reliance upon animal pollination (pollinator dependence) and its role in attracting pollinators (\protect\hyperlink{ref-ollerton2011}{Ollerton \emph{et al.} 2011}; \protect\hyperlink{ref-rodger2021}{Rodger \emph{et al.} 2021}). However, it is still unknown to what extent these different reproductive compromises determine plant-pollinator interactions.

Several studies have identified links between plant traits and plant-pollinator network properties (\protect\hyperlink{ref-bartomeus2013}{Bartomeus 2013}; \protect\hyperlink{ref-olito2015}{Olito \& Fox 2015}; \protect\hyperlink{ref-rowe2020}{Rowe \emph{et al.} 2020}). Moreover, plant traits can define species' network roles (e.g., specialists vs generalists; \protect\hyperlink{ref-lazaro2013}{Lázaro \emph{et al.} 2013}; \protect\hyperlink{ref-tur2013}{Tur \emph{et al.} 2013}). For example, plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators (\protect\hyperlink{ref-junker2013}{Junker \emph{et al.} 2013}; \protect\hyperlink{ref-coux2016}{Coux \emph{et al.} 2016}). Morphological matching between plant and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure (\protect\hyperlink{ref-stang2009}{Stang \emph{et al.} 2009}; \protect\hyperlink{ref-ibanez2012}{Ibanez 2012}). \textbf{Importantly, the combination of traits has shown to increase the predictive power of the network interactions (\protect\hyperlink{ref-eklof2013}{Eklöf \emph{et al.} 2013}) and by considering the position in the multidimensional trait space we can determine species' functional roles within plant-pollinator networks (\protect\hyperlink{ref-dehling2016}{Dehling \emph{et al.} 2016}). Nonetheless, we know little if those patterns generally studied at the community level are representative of wider macroecological scales.}

@@

All species names, genera, families and orders were retrieved and standardized from the taxonomy data sources NCBI (\url{https://www.ncbi.nlm.nih.gov/taxonomy}) for plants and ITIS (\url{https://www.itis.gov/}) for pollinators, using the R package \emph{taxize} (\protect\hyperlink{ref-chamberlain2020}{Chamberlain \emph{et al.} 2020}). We filled the `not found' searches manually using \url{http://www.theplantlist.org/} and \url{http://www.mobot.org/} for plants and \url{http://www.catalogueoflife.org/} for floral visitors.

@@

We selected a total of 19 different functional traits that included both reproductive and vegetative traits (see Table 1 and Supplementary Information). From these, 16 were reproductive traits (13 floral and 3 reproductive biology traits) and were selected based on their relevance to plant reproduction and data availability. The 3 remaining traits, were vegetative traits that are commonly used to characterize the global spectrum of plant form and function and represent the fast-slow continuum of trait variation (e.g., short-lived versus perennial species). For each plant species, we undertook an extensive literature and online search across a wide range of resources (plant databases, online floras, books, journals and images). From a total of 30,120 cells (20 columns × 1,506 species) we were able to fill 24,341 cells (80.8\% of the dataset, see Fig. S1 for missing values information for each trait).

@@

\caption{\label{tab:unnamed-chunk-1}\textbf{Table 1 | Quantitative and categorical traits used in this study.}}
\centering
\fontsize{10}{12}\selectfont
\begin{tabular}[t]{>{}cl>{}ccl}
\toprule
\multicolumn{2}{c}{\textbf{Quantitative traits}} & \multicolumn{3}{c}{\textbf{Categorical traits}} \\
\cmidrule(l{3pt}r{3pt}){1-2} \cmidrule(l{3pt}r{3pt}){3-5}
\textbf{Type} & \textbf{Traits} & \textbf{Type} & \textbf{Traits} & \textbf{Categories}\\
\midrule
\textbf{Vegetative} & Plant height (m) & \textbf{Vegetative} & Lifepan & \makecell[l]{Short-lived \\ Perennial}\\
\addlinespace
\textbf{Floral} & Flower width (mm) & \textbf{Vegetative} & Life form & \makecell[l]{Herb \\ Shrub \\ Tree}\\
\addlinespace
\textbf{Floral} & Flower length (mm) & \textbf{Floral} & Flower shape & \makecell[l]{Brush \\ Campanulate \\ Capitulum \\ Open \\ Papilionaceous \\ Tube}\\
\addlinespace
\textbf{Floral} & Inflorescence width (mm) & \textbf{Floral} & Flower symmetry & \makecell[l]{Actinomorphic \\ Zygomorphic}\\
\addlinespace
\textbf{Floral} & Style length (mm) & \textbf{Floral} & Nectar & \makecell[l]{Presence \\ Absence}\\
\addlinespace
\textbf{Floral} & Ovules per flower & \textbf{\makecell[c]{Reproductive \\ biology}} & Compatibility system & \makecell[l]{Self-incomp. \\ Part. self-comp. \\ Self-comp.}\\
\addlinespace
\textbf{Floral} & Flowers per plant & \textbf{\makecell[c]{Reproductive \\ biology}} & Breeding system & \makecell[l]{Hermaphrodite \\ Monoecious \\ Dioecious}\\
\addlinespace
\textbf{Floral} & Nectar ($\mu$l) & \textbf{} &  & \\
\addlinespace
\textbf{Floral} & Nectar (mg) & \textbf{} &  & \\
\addlinespace
\textbf{Floral} & Nectar concentration ($\%$) & \textbf{} &  & \\
\addlinespace
\textbf{Floral} & Pollen grains per flower & \textbf{} &  & \\
\addlinespace
\textbf{\makecell[c]{Reproductive \\ biology}} & Autonomous selfing (fruit set) & \textbf{} &  & \\
\bottomrule
\end{tabular}
\end{table}

@@

Trait missing values were imputed with the function \emph{missForest} (\protect\hyperlink{ref-stekhoven2012}{Stekhoven \& Bühlmann 2012}) which allows imputation of data sets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of \emph{missForest} (\protect\hyperlink{ref-penone2014}{Penone \emph{et al.} 2014}). To extract the eigenvectors, we used the function \emph{PVRdecomp} from the package \emph{PVR} (\protect\hyperlink{ref-santos2018}{Chamberlain \emph{et al.} 2018}) based on a previous conceptual framework that considers phylogenetic eigenvectors (\protect\hyperlink{ref-diniz-filho2012}{Diniz-Filho \emph{et al.} 2012}). We conducted two different imputations, one for the full set of species (1,506 species, 5.79\% of missing values) excluding nectar and pollen traits because of the high percentage of missing values (Fig. S1) and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species, 8.01\% of missing values).

@@

We explored the trade-offs between different quantitative plant functional traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson's correlation = 0.72, \emph{P} \textless{} 0.01 and Pearson's correlation = 0.36, \emph{P} \textless{} 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the association of the different qualitative traits with the main axes of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers, we excluded values within the 2.5th--97.5th percentile range (\protect\hyperlink{ref-legendre2012}{Legendre \& Legendre 2012}), and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range. We performed the pPCA using the function \emph{phyl.pca} from the package \emph{phytools} (\protect\hyperlink{ref-revell2012}{Revell 2012}) with the method lambda (\(\lambda\)) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation (\protect\hyperlink{ref-abdi2010}{Abdi \& Williams 2010}). Moreover, to corroborate that our imputation of missing values did not affect our results, we conducted a pPCA on the full dataset without missing values (see Fig. S2). We found little difference between the explained variance with the imputed dataset (51.08\%) and the dataset without missing values (52.87\%). In addition, the loadings on each principal component had a similar contribution and correlation patterns, with the exception of plant height which showed slight variations between the imputed and non-imputed dataset. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

\textbf{Phylogenetic signal}

We calculated the phylogenetic signal of the different quantitative traits on the imputed dataset with the full set of species (N = 1,506) with the package \emph{phytools} (\protect\hyperlink{ref-revell2012}{Revell 2012}) and we used Pagel's \(\lambda\) as a measurement of the phylogenetic signal. However, for pollen and nectar traits, phylogenetic signal was calculated only on the subset of species that had quantitative information for these traits (N = 755).

@@

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. These networks were analysed in their qualitative and quantitative form. First, we analysed the binary version of these weighted networks with presence-absence information that assumes equal weight across interactions. Second, we analysed the untransformed weighted networks with interaction frequency that accounts for the intensity of the interaction. Although floral visitors are not always pollinators and interaction frequency does not consider each pollinator species efficiency (\protect\hyperlink{ref-ballantyne2015}{Ballantyne \emph{et al.} 2015}), interaction frequency can provide valuable information of the contribution of floral visitors to pollination (\protect\hyperlink{ref-vazquez2005}{Vázquez \emph{et al.} 2005}, \protect\hyperlink{ref-vazquez2012}{2012}). In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across networks but this was accounted for in analyses by considering them in the random effects of the modelling process. All analyses were conducted in R version \emph{4.0.3}.

@@

We used Bayesian modelling (see below for details) to explore the effect of floral visitor groups and the main axes of trait variation (pPCA with imputed dataset) on both qualitative and quantitative floral interactions per plant species. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits, we also explored the presence-absence of interaction and visitation rate of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that \emph{Apis mellifera} was the floral visitor with the largest proportion of records counted (7.55\% of the total). This finding is consistent with previous research showing that \emph{A. mellifera} was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems (\protect\hyperlink{ref-hung2018}{Hung \emph{et al.} 2018}). Hence, to control for the effect of \emph{A. mellifera} on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of interaction and visitation rate excluding \emph{A. mellifera}. We found that \emph{A. mellifera}, was partly driving some of the observed trends on PC1 (Fig. S3). However, we did not detect major differences on PC2 and PC3.

We implemented Bayesian generalized linear mixed models using the R package \emph{brms} (\protect\hyperlink{ref-burkner2017}{Bürkner 2017}). We modelled the frequency of visits as a function of the main axes of plant trait variation and their interactions with floral visitor functional groups (Visits \textasciitilde{} PC1 x FGs + PC2 x FGs + PC3 x FGs). Because we were interested in possible differences in the visitation patterns among floral visitors groups to plants with different strategies, we included interactions between the main axes of trait variation (PC1, PC2 and PC3) and the floral visitor guilds. In this model, we added a nested random effect of networks nested within the study system to capture the variation in networks among studies and within networks. Moreover, we included the phylogenetic covariance matrix as a random factor due to the possible shared evolutionary histories of species and therefore lack of independence across them. We specified this model with a zero inflated negative binomial distribution and weakly informative priors from the brms function. We run this model for 3,000 iterations and with previous 1,000 warm up iterations. We set delta (\(\Delta\)) to 0.99 to avoid divergent transitions and visualized the posterior predictive checks with the function \emph{pp\_check} using the \emph{bayesplot} package (\protect\hyperlink{ref-gabry2019}{Gabry \emph{et al.} 2019}).
